Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-08-19 main.tex
Updates from Overleaf
began reviewer 1 revisions

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
% my added packages and commands =======================
\usepackage[numbers]{natbib} %removed cite package
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}         
\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
\providecommand{\gls}[1]{#1}
\definecolor{highlight}{RGB}{00,150,00}
\definecolor{todo}{RGB}{200,50,00}
\definecolor{rank1}{HTML}{70FF70}
\definecolor{rank2}{HTML}{858585}
\definecolor{rank3}{HTML}{454545}
\definecolor{rank4}{HTML}{000000}
\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
\newcommand{\JTFS}{\texttt{JTFS}}
\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
\newcommand{\OutDomain}{\textbf{Out-of-Domain Generation}}
\newcommand{\BPNoise}{\textbf{BP-Noise}}  
\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
\newcommand{\AmpMod}{\textbf{Noise-AM}}  
\newcommand{\FMMod}{\textbf{SineSaw-AM}}

@@

\section{Introduction}
\subsection{Digital Synthesis With Sound-Matching}
Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer. Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation. However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds. One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods. As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.

@@

A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.

@@

\subsection{Automatic Sound Matching}
\label{sec:sound_matching_definition}
In past works the sound-matching problem has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$. To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
\begin{itemize}
    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
    \item $t$: The target sound which we want to replicate or imitate. 
    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
\end{itemize}

It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs. Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.

@@

\subsubsection{Determining Best Performers}
 Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked. We have four loss functions and four synthesizer programs. The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination. Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS. Two of the authors assigned blinded Likert scores to 40 randomly selected sounds, where integers from 1 to 5 are assigned, with 1 meaning no similarity and 5 for near identical.

@@

\subsection{Loss Function Implementation Details}
\label{sec:loss_implementation}
\subsubsection{STFT Losses}
Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two of the loss functions. As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings. An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies. If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity. In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison. In the computer vision field, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.

>>>>

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
% my added packages and commands =======================
\usepackage[numbers]{natbib} %removed cite package
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}

\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
\providecommand{\gls}[1]{#1}
\definecolor{rank1}{HTML}{70FF70}
\definecolor{rank2}{HTML}{858585}
\definecolor{rank3}{HTML}{454545}
\definecolor{rank4}{HTML}{000000}
\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
\newcommand{\JTFS}{\texttt{JTFS}}
\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
\newcommand{\BPNoise}{\textbf{BP-Noise}}  
\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
\newcommand{\AmpMod}{\textbf{Noise-AM}}  
\newcommand{\FMMod}{\textbf{SineSaw-AM}}

@@

\section{Introduction}
Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer. Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation. However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds. One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods. As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.

@@

\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}. }

@@

\subsection{Automatic Sound Matching}
\label{sec:sound_matching_definition}
The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$. To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
\begin{itemize}
    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
    \item $t$: The target sound which we want to replicate or imitate. 
    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
\end{itemize}

\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}. Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.

@@

\subsubsection{Determining Best Performers}
 Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked. We have four loss functions and four synthesizer programs. The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination. Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.

@@

\subsection{Loss Function Implementation Details}
\label{sec:loss_implementation}
\subsubsection{STFT Losses}
Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions. As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings. An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies. If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a linear equation $f(x) = ax$ can never overlap with a slightly shifted version of itself, or $f(x) = ax +\epsilon$). In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison. In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
